---
title: "Rank Order Scatter Plot"
output: html_document
date: "2023-01-19"
---

Load packages:
```{r}
library(ggplot2)
library(ggExtra)
library(readr)
library(ggplotgui)
library(plotly)
```

Show full current path:
```{r}
getwd( )
```

# CorPGA Cac Scatterplot 

Read the csv table:
```{r}
Cac_size <- read_csv("Cac_size.csv")
View(Cac_size)
```

Reorder from smallest to largest genome size (bp):
```{r}
Cac_Size_order <- Cac_size[order(Cac_size$Size),]
```

Show new rearranged table based on size (bp):
```{r}
print(Cac_Size_order)
```

Save the file as csv:
```{r}
write.csv(Cac_Size_order, "Cac_size_order.csv", row.names=FALSE)
```

Make dataframe and view:
```{r}
Cac_size_order <- read_csv("Cac_size_order.csv")
View(Cac_size_order)
```

Create rank order scatter plot Cac:
```{r}
graph <- ggplot(Cac_size_order) +
  aes(x = reorder(Genome, Size), y = Size) +
  geom_point() +
  scale_x_discrete(name="Genome", guide = guide_axis(angle = 45)) +
  scale_y_continuous(name="Size (bp)", limits=c(2350000, 2650000))
```

```{r}
ggsave('Cac_size_order_scatter.png', graph, width = 14, height = 14, units = 'cm')
```

Save scatter plot: 
```{r}
ggsave('Cac_size_order_scatter.png', graph)
```

Load plotly and use ggplotly on ggplot:
```{r}
ggplotly(graph, height = 500, width=800)
```

Save plotly as html:
```{r}
l <- plotly::ggplotly(graph)

htmlwidgets::saveWidget(l, "Cac_size_order.html", selfcontained = TRUE)
```


# CorPGA Cps Scatterplot

```{r}
Cps_size_order <- read_csv("Cps_size.csv")
View(Cps_size_order)
```

```{r}
Cps_Size_order <- Cps_size_order[order(Cps_size_order$Size),]
```

Show new rearranged table based on size (bp):
```{r}
print(Cps_Size_order)
```

```{r}
graph <- ggplot(Cps_size_order) +
  aes(x = reorder(Genome, Size), y = Size) +
  geom_jitter() +
  scale_x_discrete(name="Genome", guide = guide_axis(angle = 45)) +
  scale_y_continuous(name="Size (bp)", limits=c(2250000, 2450000))
```

```{r}
ggsave('Cps_size_order_scatter.png', graph)
```

# Cps rank order genome size (control)

```{r}
Cpr_size_order <- read_csv("Cpr_size.csv")
View(Cpr_size_order)
```

```{r}
Cpr_Size_order <- Cpr_size_order[order(Cpr_size_order$Size),]
print(Cpr_Size_order)
```

```{r}
graph <- ggplot(Cpr_size_order) +
  aes(x = reorder(Genome, Size), y = Size) +
  geom_jitter() +
  scale_x_discrete(name="Genome", guide = guide_axis(angle = 45)) +
  scale_y_continuous(name="Size (bp)", limits=c(2420000, 2600000))
```

```{r}
ggsave('Cpr_size_order_scatter.png', graph)
```

# NovCor

Cna and feet clade